Calculating Entropies With Nested Sampling

### 2017-02-27 paper/ms.tex
talking about the cost

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{\theta}}
\newcommand{\n}{\boldsymbol{\eta}}

@@

The computational resources needed to compute these quantities are quite large,
though the algorithm can be parallelised trivially. However, the problem
being solved is not trivial. The mutual information of $\tau$ and
$\d$ is
\begin{align}
I_{\tau, \d} &= \int p(\x, \d) \log\frac{p(\x, \d)}{p(\x)p(\d)} \, d\x \, d\d.
\end{align}
However, $\x$ is one of several parameters. The others, $A$, and $\phi$,
were implicitly integrated out. Denoting the nuisance parameters by
$\n$, the mutual information that has been calculated is
\begin{align}
I_{\x, \d} &= \int p(\x, \d)
  \log\frac{\int p(\x, \n, \d) \, d\n }
           {p(\x) \int p(\x', \n)p(\d | \x', \n) \, d\x' \, d\n}
  \, d\x \, d\d.
\end{align}
It should not be considered surprising that this is a challenging task.

\subsection{Uneven observations}
The results for the uneven observing schedule are as follows:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 106.8 \pm 1.1   \\
H'_{\tau, \d} &= 97.0 \pm 1.0    \\
I_{\tau, \d} &= 10.9 \pm 0.4
\end{align}
which appears to be higher than under the even observing schedule,
though the uncertainties
are not small enough to be certain of this.

@@

>>>>

\renewcommand{\d}{\boldsymbol{d}}
\newcommand{\todo}{\color{orange} \bf}
\newcommand{\query}{\color{green} \bf}
\newcommand{\x}{\boldsymbol{\theta}}
\newcommand{\bphi}{\boldsymbol{\phi}}
\newcommand{\boldeta}{\boldsymbol{\eta}}
\newcommand{\n}{\boldsymbol{\eta}}

@@

\subsection{Uneven observations}

The results for the uneven observing schedule are as follows:
\begin{align}
H'_{\tau}     &= 1.122 \pm 0.028 \\
H'_{\d}       &= 106.8 \pm 1.1   \\
H'_{\tau, \d} &= 97.0 \pm 1.0    \\
I_{\tau, \d} &= 10.9 \pm 0.4
\end{align}
which appears to be higher than under the even observing schedule,
though the uncertainties
are not small enough to be certain of this.

@@

\subsection{Computational cost}

The computational resources needed to compute these quantities was quite large
--- three lots of 1,000 repetitions of Nested Sampling, in all
(though the algorithm can be parallelised trivially).
However, it is worth reflecting on the difficulty of the calculation that
has been done.

In a Bayesian inference problem with parameters $\x$ and data $\d$, the
mutual information is
\begin{align}
I_{\x, \d} &= \iint p(\x, \d)
                        \ln \left[\frac{p(\x, \d)}{p(\x)p(\d)}\right]
                        \, d\x \, d\d
\end{align}
and this measures the dependence between $\x$ and $\d$. A Monte Carlo
strategy to evaluate this is to sample from $p(\x, \d)$ and average
the value of the logarithm, which involves computing the log-evidence
\begin{align}
\ln p(\d) = \ln\left[\int p(\x)p(\d | \x) \, d\x\right]
\end{align}
for each possible data set. Computing $p(\d)$ for even a single
data set has historically been considered difficult.

However, if we want the mutual information between the data and
{\em a subset of the parameters} (i.e., there are nuisance parameters
we don't care about), things become even more tricky. Suppose we
partition the parameters $\x$ into important parameters
$\bphi$ and nuisance parameters $\boldeta$, such that we want to calculate
the mutual information $I_{\bphi, \d}$. This is given by
\begin{align}
I_{\bphi, \d} &= \iint p(\bphi, \d)
                        \ln \left[\frac{p(\bphi, \d)}{p(\bphi)p(\d)}\right]
                        \, d\bphi \, d\d
\end{align}
